Efficient and Scalable Bipartite Matching with Fast Beta Linkage (fabl)

### 2024-01-27 ba_resubmit_template_files/response_to_reviewers4-rcs-revision.tex

	\title{Response to Reviewer Comments}
	\maketitle

We thank the Editor and Associate Editor for their assessment and constructive feedback.

\section{Editor Comments}

\pointRaised{E}{I received a detailed and careful report from the original associate editor who thinks the paper is now easier to read and understand. I still find it intricate to read, but the AE may have a point when saying that the notation can be a difficult challenge in this area of research.}
\reply{}

\pointRaised{E}{Anyhow, I have decided to follow the AE’s recommendation and would like the authors to carefully consider the comments in the AE report and make the suggested revisions. I find the issue raised by the AE regarding the formulas in Section 4.2 particularly concerning for two reasons. First, the article cannot be accepted if it contains technical errors and inaccuracies. Second, the formulas in question were used to code up the samplers used in the simulations and data analysis examples. If the formulas are incorrect (do not make sense, in fact, as the AE points out) how can the computational results be trusted? This is a grave concern that must be satisfactorily and convincingly addressed before the manuscript is accepted for publication.}
\reply{}

\pointRaised{E}{Please prepare a revision addressing all the AE’s comments, explain how the errors in the formulas affected the computations, and revise the computations as needed.}
\reply{}
%Thank you for the thoughtful feedback. We have made several broad changes to address the Editor's and Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.

\section{Associate Editor Comments}

\textcolor{red}{Brian: change E to AE. Make sure to address the reviewer as AE (many typos below). Address the questions/comments of the Editor above.}

@@

Despite the confusion about this specific sentence, the proposed change in notation more explicitly shows how the sampler handles non-matches, which we think is the editors primary concern. 
}

\pointRaised{E}{Page 9, just below equation (17): Should the reference to equation (9) should be updated to equation (13)?}

\reply{The reference to equation (9) was meant as a comparison with equations (18) and (19), not equation (17). We have revised the paragraph to make this more clear. 
}

\textcolor{red}{Could you add in the new paragraph so it's more clear to the AE and for us to review please?}

@@

\pointRaised{E}{Supplement A, equation (12): I believe that this intermediate result is not described correctly. Because $c_j$ is a function of $\gamma_{ij}$, once the function is divided by $c_j$, the quantity is no longer proportional to the probability. Rather, it is proportional to a likelihood-like function. Because equation (13) is conditional on $\gamma$, this distinction is not important for the end result.
}

\reply{Since $\Gamma_{.j}$ is the data, the expression $p(\Gamma_{.j} | -)$ is intended to be interpreted as a likelihood. To avoid this confusion, and to be consistent with notation in equation (1) of the Supplement, and equations (15) and (17) in the main paper, we have revised equations (8) - (12) of the supplement to use $\mathcal{L}(Z_j|-)$ instead of $p(\Gamma_{.j}|-)$.

{
\color{red}{I added the changed notation in red. However, it relies on fairly ugly (and I think non-standard) uses of likelihood notation (for example, $\mathcal{L}(Z_j = q | -)$). In contrast, Jody's paper stayed with $p(\Gamma_{.j} | -)$ notation. Should we make the switch, or stand by the old notation? The reviewer acknowledged that it practically doesn't make a difference, so I'm perhaps inclined to stick with the version we submitted?
}
}
}

\textcolor{red}{I believe the AE is stating that things should be equality and no proportionality. If you look at what he/she wrote up this seems to be what he/she is stating to be the case. In essence, we cannot drop the constants. Please let me know if you agree regarding this Jerry as it's worded a bit strangely to me.}

### 2024-01-27 ba_resubmit_template_files/response_to_reviewers4-rcs-revision.tex

\pointRaised{E}{Anyhow, I have decided to follow the AE’s recommendation and would like the authors to carefully consider the comments in the AE report and make the suggested revisions. I find the issue raised by the AE regarding the formulas in Section 4.2 particularly concerning for two reasons. First, the article cannot be accepted if it contains technical errors and inaccuracies. Second, the formulas in question were used to code up the samplers used in the simulations and data analysis examples. If the formulas are incorrect (do not make sense, in fact, as the AE points out) how can the computational results be trusted? This is a grave concern that must be satisfactorily and convincingly addressed before the manuscript is accepted for publication.}
\reply{}

\pointRaised{E}{Please prepare a revision addressing all the AE’s comments, explain how the errors in the formulas affected the computations, and revise the computations as needed.}
\reply{}
%Thank you for the thoughtful feedback. We have made several broad changes to address the Editor's and Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.

\section{Associate Editor Comments}

\textcolor{red}{Brian: change E to AE. Make sure to address the reviewer as AE (many typos below). Address the questions/comments of the Editor above. Please separate out major comments with minor comments as the AE did.}

\pointRaised{AE}{For Section 4.2, my question starts with the definition of $\Gamma_{Z_{j^{(s+1)}, j}}.$ Is this quantity always defined for possible values of $Z$? That is, if $Z_{j}^{(s+1)} = n_A + j$, what is the meaning of $\Gamma_{n_A+j, j}$?}

### 2024-01-27 ba_resubmit_template_files/response_to_reviewers4.tex

We thank the Editor and Associate Editor for their assessment and constructive feedback.

%Thank you for the thoughtful feedback. We have made several broad changes to address the Editor's and Associate Editor's overall critiques, and have addressed each specific comment from the reviewer.

\section{Editor Comments}

%\pointRaised{E}{I received a detailed and careful report from the original associate editor who thinks the paper is now easier to read and understand. I still find it intricate to read, but the AE may have a point when saying that the notation can be a difficult challenge in this area of research.}
%\reply{}

\pointRaised{E}{Anyhow, I have decided to follow the AE’s recommendation and would like the authors to carefully consider the comments in the AE report and make the suggested revisions. I find the issue raised by the AE regarding the formulas in Section 4.2 particularly concerning for two reasons. First, the article cannot be accepted if it contains technical errors and inaccuracies. Second, the formulas in question were used to code up the samplers used in the simulations and data analysis examples. If the formulas are incorrect (do not make sense, in fact, as the AE points out) how can the computational results be trusted? This is a grave concern that must be satisfactorily and convincingly addressed before the manuscript is accepted for publication.}

\reply{The Associate Editor made thoughtful suggestions for revising the notation in Section 4.2. We have adopted these suggestions as discussed below. We assure the Editor that formulas used in all computations is correct, and had already reflected the corrected notation in the previous submission.}

\section{Associate Editor Comments}

\pointRaised{AE}{For Section 4.2, my question starts with the definition of $\Gamma_{Z_{j^{(s+1)}, j}}.$ Is this quantity always defined for possible values of $Z$? That is, if $Z_{j}^{(s+1)} = n_A + j$, what is the meaning of $\Gamma_{n_A+j, j}$?}

\reply{The Associate Editor is correct that $\Gamma_{n_A+j, j}$ is not defined under this notation. We had attempted to remove all instances in which this notation might be implied, but it seems we did not catch them all. We thank the AE for their thoughtful suggestions in this regard.
}

\pointRaised{AE}{This notation first arises in the definition of $n_p(Z)$, which I think could easily be changed to something along the lines of $n_p(Z) = \sum_{j:Z_j \leq n_A} I(\gamma_{Z_j, j} = h_p)$.
}

@@

\reply{Thank you for this suggestion. We have adopted this change.
}

\pointRaised{AE}{My bigger concern comes for equation (18), where I am not entirely sure what probability is being computed, or how this sampling procedure combined with equation (19) would result in a “non-match” where $Z_j = n_A + j$. I believe that the idea is to partition the possible values of $Z_j:\left( \{r_{p_j}\}, n_A + j\right)$. Then, in equation (18), the left hand side is intended to be the probability that Z falls into each part of the partition. But, in the revision, I didn’t see where the last “non-match” part of the partition is defined, and so there is no real way to select the non-match from equation (19).
}

\reply{The AE's understanding of the sampler is correct. The AE also correctly identified an error in equation (19), such that there was no real way to select the non-match option. We thank the AE for the recommended change in notation, and have adopted the recommendation. The sampling equations are now provided as follows:

Define $r$ as an arbitrary set of records. We have
\begin{align}
	\label{eqn:gibbs1}
	p\left( Z_j^{(s+1)} \in r \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
	\begin{cases} 
		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & r = r_{p_j}; \\
		1- \pi^{(s+1)} , &  r = \{n_A + j\}. \\
	\end{cases}
\end{align}
Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
\begin{align}
	\label{eqn:gibbs2}
	p\left(Z_j^{(s+1)} = q \mid Z_j^{(s+1)} \in r, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
		\frac{1}{N_{p_j}}, & r = r_{p_j} \text{ and } q \in r; \\
		1, & r = \{n_A + j\} \text{ and} \\
		& q = n_A + j. \\
	\end{cases}
\end{align} 
}

\pointRaised{AE}{Also, it’s not clear to me as currently written that the authors are equating the patterns with that partition so that by definition a “true” match that does not appear in A cannot have a pattern that appears in $\mathcal{P}$, but did not happen to be observed (because the match was not captured in $A$).
}

@@

\reply{We did not quite understand this part of the comment. If record $B_j$ has no match in $A$, then $Z_j = n_A + j$. If there is no matching record, there is no comparison vector (or pattern) for such a non-existent record.

Despite the confusion about this specific sentence, the proposed change in notation more explicitly shows how the sampler handles non-matches, which we think is the AE's primary concern. 
}

\pointRaised{AE}{Page 9, just below equation (17): Should the reference to equation (9) should be updated to equation (13)?}

\reply{The reference to equation (9) was meant as a comparison with equations (18) and (19), not equation (17). We have revised the paragraph to make this more clear. This material now reads:

Note that sampling from (18) has complexity $O(P+1)$ and sampling (uniformly) from (19) has complexity $O(1)$, regardless of the size of $r_{p_j}$. In contrast, sampling $Z_j$ from the full conditional provided in (9) has complexity $O(n_A)$, because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities. 
}

\pointRaised{AE}{Supplement A, just below equation (3): Should the summation be over index $j$?
}

@@

\reply{Yes. We have made the correction. 
}

\pointRaised{AE}{Supplement A, equation (12): I believe that this intermediate result is not described correctly. Because $c_j$ is a function of $\gamma_{ij}$, once the function is divided by $c_j$, the quantity is no longer proportional to the probability. Rather, it is proportional to a likelihood-like function. Because equation (13) is conditional on $\gamma$, this distinction is not important for the end result.
}

\reply{We have removed the intermediate result in equation (12).}

### 2024-01-27 ba_resubmit_template_files/main4.tex
bk eddits

The posterior updates for $\bm{m}$ and $\bm{u}$ depend on the data only through quantities that can be calculated through $\mathcal{N}$ and $\mathcal{P}$. Let $n_p(\bm{Z}) = \sum_{j: Z_j \leq n_A} I\left(\gamma_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ be the total occurrence of pattern $p$ in the data across all record pairs. Then, conditional on $\bm{Z}$, we can express the contribution to the likelihood in the full conditional for $\bm{m}$ and $\bm{u}$ as
\begin{align}
	\mathcal{L}(\bm{m}, \bm{u} \mid \tilde{\gamma}, \bm{Z}, \pi)  &=  \prod_{p=1}^P m_p^{n_p(\bm{Z})}u_p^{N_p - n_p(\bm{Z})}  \label{likelihood_efficient_m_u}.
\end{align}
Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ and $\bm{\beta_0}= (\beta_{11}, \ldots, \beta_{F L_F})$ be concatenated vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively. The terms needed for the posterior updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
\begin{subequations}
	\begin{align}
		\bm{\alpha(Z}^{(s)}\bm{)} &= \bm{\alpha_0} + \sum_{p=1}^P n_p\left(\bm{Z}^{(s)}\right) \times e(h_p), \label{efficient_alpha} \\
		\bm{\beta(Z}^{(s)}\bm{)} &= \bm{\beta_0} + \sum_{p=1}^P \left(N_p - n_p\left(\bm{Z}^{(s)}\right)\right) \times e(h_p). \label{efficient_beta}
	\end{align}
\end{subequations}
Specifically, the $l + \sum_{k=1}^{f-1} L_k$ components of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provide the posterior updates for level $l$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}). However, the vectorized summations can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
{
\color{blue}{
Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$. For each pattern $p$, let $w_p =m_p/u_p$.  The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
\begin{align}
	\mathcal{L}(Z_j\mid  \tilde{\gamma}, \bm{m}, \bm{u}, \pi) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}. \label{likelihood_efficient_z}
\end{align}
Sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) can become computationally expensive when $n_A$ is large. This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$. To speed up computation, we use \eqref{likelihood_efficient_z} to break this sampling step into two.

We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
\begin{align}
	%\label{eqn:gibbs1}
	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
	\begin{cases} 
		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & h_p \in \mathcal{P}; \\
		1- \pi^{(s+1)} , &   \text{otherwise}. \\
	\end{cases}
\end{align}
Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
\begin{align}
	%\label{eqn:gibbs2}
	p\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
		0, & \text{otherwise.} \\
	\end{cases}
\end{align}

These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups. We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.
}
}

{
\color{red}{Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$. For each pattern $p$, let $w_p =m_p/u_p$.  The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
\begin{align}
	\mathcal{L}(Z_j\mid  \tilde{\gamma}, \bm{m}, \bm{u}, \pi) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}. \label{likelihood_efficient_z}
\end{align}
With this form, we break this sampling step into two. We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
\begin{align}
	\label{eqn:gibbs1}
	p\left( Z_j^{(s+1)} \in r \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
	\begin{cases} 
		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & r = r_{p_j}; \\
		1- \pi^{(s+1)} , &   r = n_A + j. \\
	\end{cases}
\end{align}
Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
\begin{align}
	\label{eqn:gibbs2}
	p\left(Z_j^{(s+1)} = q \mid Z_j^{(s+1)} \in r, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
		\frac{1}{N_{p_j}}, & r = r_{p_j} \text{ and } q \in r; \\
		1, & r = n_A + j. \\
	\end{cases}
\end{align} 
Note that sampling from (\ref{eqn:gibbs1}) has complexity $O(P+1)$ and sampling (uniformly) from (\ref{eqn:gibbs2}) has complexity $O(1)$, regardless of the size of $r_{p_j}$. In contrast, sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) has complexity $O(n_A)$, because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities.

These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups. We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.}
}
\begin{lemma}
	Let $n_A$ and $n_B$ be the number of records in files $A$ and $B$, respectively. Let $F$ be the number of fields used for comparisons across records, and $P$ be the number of patterns that comparison vectors exhibit in $A \times B$. We assume $C$ cores available for parallelization and a Gibbs sampler with $T$ iterations. Then, the overall computational complexity of \texttt{fabl} with hashing is $O(F n_A n_B /C) + O(T n_B P / C)$.
	\label{lemma:fabl}
\end{lemma}
\begin{proof}
	We consider two steps: constructing the comparison vectors and the Gibbs sampler. The computational complexity of all pairwise comparisons across $A$ and $B$ is $O(F n_A n_B)$. The hashing procedure for all pairwise comparisons is also $O(F n_A n_B)$. With $C$ processors available, we can split these computations across $C$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(F n_A n_B /C)$. There are then trivial computational costs associated with synthesizing summary statistics across these partitions. We note that this contribution to computational complexity applies for all versions of \cite{fellegi_theory_1969} algorithms unless they use blocking to reduce the comparison space.

@@

		We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$, and delete the memory intensive matrix $\gamma^{ab}$ before continuing with the next batch of data. In detail, we calculate
\begin{subequations}
	\begin{align}
		r_{p_j}^{ab} &= \{i \in 1, \dots, n_A| \gamma_{ij} = h_p, B_j \in B^b\}, \label{r_{p_j}^{ab}} \\
		N_{p_j}^{ab} &= |r_{p_j}^{ab}|. \label{N_{p_j}^{ab}}
	\end{align}
\end{subequations}
These can be computed sequentially or in parallel. Summary statistics from each pairwise batch comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
\begin{subequations}
	\begin{align}
		r_{p_j} &= (r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B}) \text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B, \label{r_{p_j}} \\
		N_{p_j} &= \sum_{a = 1}^{t_A} \sum_{b = 1}^{t_B} N_{p_j}^{ab}.\label{N_{p_j}}	\end{align}
\end{subequations}

@@

All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$. These records have the same probability to be identified as the link for record $B_j$. Thus, records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate. Rather than store all of these record labels, we store only a small number $S$. For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form SEI$(r_{p_j}^{ab})$. We collect these memory-reduced lists to form SEI$(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form SEI$(\mathcal{R})$. 
{
\color{blue}{
As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$. We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
}
}
{
\color{red}{As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), and (\ref{eqn:gibbs1}), all original record pairs are still accounted for through $\mathcal{N}$. We must revise (\ref{eqn:gibbs2}) to account for the reduced number of labels, so instead use
\begin{align}
	\label{eqn:gibbs2-SEI}
	p\left(Z_j^{(s+1)} = q \mid Z_j^{(s+1)} \in r, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
		\frac{1}{|r|}, & r = \text{SEI}(r_{p_j}) \text{ and} \\
		& q \in r; \\
		1, & r = n_A + j \\
	\end{cases}
\end{align}
in the second step of the split sampler. In this way, we can conduct all posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$. We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
}
}

>>>>

The posterior updates for $\bm{m}$ and $\bm{u}$ depend on the data only through quantities that can be calculated through $\mathcal{N}$ and $\mathcal{P}$. Let $n_p(\bm{Z}) = \sum_{j: Z_j \leq n_A} I\left(\gamma_{Z_j, j} = h_p \right)$ be the number of matching record pairs with agreement pattern $p$, and $N_p = \sum_{j=1}^{n_B} N_{p_j}$ be the total occurrence of pattern $p$ in the data across all record pairs. Then, conditional on $\bm{Z}$, we can express the contribution to the likelihood in the full conditional for $\bm{m}$ and $\bm{u}$ as
\begin{align}
	\mathcal{L}(\bm{m}, \bm{u} \mid \tilde{\gamma}, \bm{Z}, \pi)  &=  \prod_{p=1}^P m_p^{n_p(\bm{Z})}u_p^{N_p - n_p(\bm{Z})}  \label{likelihood_efficient_m_u}.
\end{align}
Additionally, let $\bm{\alpha_0} = (\alpha_{11}, \ldots, \alpha_{F L_F})$ and $\bm{\beta_0}= (\beta_{11}, \ldots, \beta_{F L_F})$ be concatenated vectors of prior parameters for the $\bm{m}$ and $\bm{u}$ distributions respectively. The terms needed for the posterior updates for the $\bm{m}$ and $\bm{u}$ parameters are given by the appropriate components of the vectors
\begin{subequations}
	\begin{align}
		\bm{\alpha(Z}^{(s)}\bm{)} &= \bm{\alpha_0} + \sum_{p=1}^P n_p\left(\bm{Z}^{(s)}\right) \times e(h_p), \label{efficient_alpha} \\
		\bm{\beta(Z}^{(s)}\bm{)} &= \bm{\beta_0} + \sum_{p=1}^P \left(N_p - n_p\left(\bm{Z}^{(s)}\right)\right) \times e(h_p). \label{efficient_beta}
	\end{align}
\end{subequations}
Specifically, the $l + \sum_{k=1}^{f-1} L_k$ components of (\ref{efficient_alpha}) and (\ref{efficient_beta}) provide the posterior updates for level $l$ and field $f$ in (\ref{eqn:alpha_update}) and (\ref{eqn:beta_update}). However, the vectorized summations can be computed more efficiently than summing over $n_A n_B$ record pairs for each field and agreement level.
%{
%\color{blue}{
%Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$. For each pattern $p$, let $w_p =m_p/u_p$.  The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
%\begin{align}
%	\mathcal{L}(Z_j\mid  \tilde{\gamma}, \bm{m}, \bm{u}, \pi) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}. \label{likelihood_efficient_z}
%\end{align}
%Sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) can become computationally expensive when $n_A$ is large. This is because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities, which has a computational cost that scales with $n_A$. To speed up computation, we use \eqref{likelihood_efficient_z} to break this sampling step into two. 
%
%We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link, according to
%\begin{align}
%	%\label{eqn:gibbs1}
%	p\left( \Gamma_{Z_j^{(s+1)}, j} = h_p \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
%	\begin{cases} 
%		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & h_p \in \mathcal{P}; \\
%		1- \pi^{(s+1)} , &   \text{otherwise}. \\
%	\end{cases}
%\end{align}
%Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
%\begin{align}
%	%\label{eqn:gibbs2}
%	p\left(Z_j^{(s+1)} = q \mid \gamma_{Z_j^{(s+1)}, j} = h_p, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
%		\frac{1}{N_{p_j}}, & q \in r_{p_j}; \\
%		0, & \text{otherwise.} \\
%	\end{cases}
%\end{align} 
%
%These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups. We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.
%}
%}

Similarly, the posterior updates for $\bm{Z}$ depend on the data only through quantities calculated through $\mathcal{N}, \mathcal{P}$, and $\mathcal{R}$. For each pattern $p$, let $w_p =m_p/u_p$.  The contribution to the likelihood in the full conditional for $Z_j$ can be expressed as
\begin{align}
	\mathcal{L}(Z_j\mid  \tilde{\gamma}, \bm{m}, \bm{u}, \pi) = \prod_{p=1}^P u_p^{N_{p_j}} \prod_{i \in r_{p_j}} w_p^{I(Z_j = i)}. \label{likelihood_efficient_z}
\end{align}
With this form, we break this sampling step into two. We first sample among $P + 1$ options for the agreement pattern between $B_j$ and its potential link. Define $r$ as an arbitrary set of records. We have
\begin{align}
	\label{eqn:gibbs1}
	p\left( Z_j^{(s+1)} \in r \mid \tilde{\gamma}, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)}\right) \propto
	\begin{cases} 
		\frac{\pi^{(s+1)}N_{p_j}}{n_A}  w_{p}^{(s+1)},  & r = r_{p_j}; \\
		1- \pi^{(s+1)} , &  r = \{n_A + j\}. \\
	\end{cases}
\end{align}
Since all records in $A$ sharing the same agreement pattern with $B_j$ are equally likely, we then sample among candidate records uniformly using
\begin{align}
	\label{eqn:gibbs2}
	p\left(Z_j^{(s+1)} = q \mid Z_j^{(s+1)} \in r, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
		\frac{1}{N_{p_j}}, & r = r_{p_j} \text{ and } q \in r; \\
		1, & r = \{n_A + j\} \text{ and} \\
		& q = n_A + j. \\
	\end{cases}
\end{align} 
Note that sampling from (\ref{eqn:gibbs1}) has complexity $O(P+1)$ and sampling (uniformly) from (\ref{eqn:gibbs2}) has complexity $O(1)$, regardless of the size of $r_{p_j}$. In contrast, sampling $Z_j$ from the full conditional provided in (\ref{eqn:z_full_conditional}) has complexity $O(n_A)$, because sampling a value from $n_A$ options with unequal weights requires normalizing the weights to probabilities.

These changes can greatly improve the speed of the sampler, and each can be parallelized if desired for additional computational speed-ups. We emphasize the computational gains of this split sampler through Lemma \ref{lemma:fabl}.

\begin{lemma}
	Let $n_A$ and $n_B$ be the number of records in files $A$ and $B$, respectively. Let $F$ be the number of fields used for comparisons across records, and $P$ be the number of patterns that comparison vectors exhibit in $A \times B$. We assume $C$ cores available for parallelization and a Gibbs sampler with $T$ iterations. Then, the overall computational complexity of \texttt{fabl} with hashing is $O(F n_A n_B /C) + O(T n_B P / C)$.
	\label{lemma:fabl}
\end{lemma}
\begin{proof}
	We consider two steps: constructing the comparison vectors and the Gibbs sampler. The computational complexity of all pairwise comparisons across $A$ and $B$ is $O(F n_A n_B)$. The hashing procedure for all pairwise comparisons is also $O(F n_A n_B)$. With $C$ processors available, we can split these computations across $C$ equally sized partitions and compute these comparisons in parallel, so the complexity becomes $O(F n_A n_B /C)$. There are then trivial computational costs associated with synthesizing summary statistics across these partitions. We note that this contribution to computational complexity applies for all versions of \cite{fellegi_theory_1969} algorithms unless they use blocking to reduce the comparison space.

@@

		We then conduct hashing, obtain the compressed $\tilde{\gamma}^{ab}$, and delete the memory intensive matrix $\gamma^{ab}$ before continuing with the next batch of data. In detail, we calculate
\begin{subequations}
	\begin{align}
		r_{p_j}^{ab} &= \{i \in 1, \dots, n_A| \gamma_{ij} = h_p, B_j \in B^b\}, \label{r_{p_j}^{ab}} \\
		N_{p_j}^{ab} &= |r_{p_j}^{ab}|. \label{N_{p_j}^{ab}}
	\end{align}
\end{subequations}
These can be computed sequentially or in parallel. Summary statistics from each pairwise batch comparison can be synthesized to recover summary statistics for the full comparison matrix $\gamma$ through
\begin{subequations}
	\begin{align}
		r_{p_j} &= \{r_{p_j}^{11}, \ldots, r_{p_j}^{ t_A t_B} \}\text{ for } a = 1, \ldots, t_A \text{ and } b = 1, \ldots, t_B, \label{r_{p_j}} \\
		N_{p_j} &= \sum_{a = 1}^{t_A} \sum_{b = 1}^{t_B} N_{p_j}^{ab}.\label{N_{p_j}}	\end{align}
\end{subequations}

@@

All records $A_i$ that share agreement pattern $p$ with record $B_j$ have the same $w_{p}$. These records have the same probability to be identified as the link for record $B_j$. Thus, records $i \in r_{p_j}$ with large $N_{p_j}$ are unlikely to be sampled consistently enough to be deemed a match in the Bayes estimate. Rather than store all of these record labels, we store only a small number $S$. For each $r_{p_j}^{ab}$, we sample $S$ indices without replacement to form SEI$(r_{p_j}^{ab})$. We collect these memory-reduced lists to form SEI$(r_{p_j})$ as in (\ref{r_{p_j}}), and collect these to form SEI$(\mathcal{R})$. 
%{
%\color{blue}{
%As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), (\ref{eqn:gibbs1}), and (\ref{eqn:gibbs2}), all original record pairs are still accounted for through $\mathcal{N}$, and thus we can proceed with posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$. We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
%}
%}

As shown in the full conditionals in (\ref{efficient_alpha}),  (\ref{efficient_beta}), and (\ref{eqn:gibbs1}), all original record pairs are still accounted for through $\mathcal{N}$. We must revise (\ref{eqn:gibbs2}) to account for the reduced number of labels, so instead use
\begin{align}
	\label{eqn:gibbs2-SEI}
	p\left(Z_j^{(s+1)} = q \mid Z_j^{(s+1)} \in r, \bm{m}^{(s+1)}, \bm{u}^{(s+1)}, \pi^{(s+1)} \right) = \begin{cases} 
		\frac{1}{|r|}, & r = \text{SEI}(r_{p_j}) \text{ and} \\
		& q \in r; \\
		1, & r = \{n_A + j\} \text{ and} \\
		& q = n_A + j \\
	\end{cases}
\end{align}
in the second step of the split sampler. In this way, we can conduct all posterior inference with the memory reduced SEI$(\tilde{\gamma}) = \{\mathcal{P}, \text{SEI}(\mathcal{R}), \mathcal{N}\}$. We provide guidance on choice of $S$ through a simulation in Section~\ref{SEI-sensitivity}.
